draft 1

### 2016-06-14 finalwriteup/draft.rmd
started review of draft

****

@@

Predicting how long it will take a person to do something is a notoriously tricky task. People make mental calculations of these sort every day, guessing how long it will take to drive to meet someone on time, how many tasks you can check off in a day, or how long it will take to cook a dish. These mental calculations can be particularly inaccurate if the individual has never performed the task before. These errors are trivial in everyday activities, but they can have more meaning for businesses that sell their time as their chief product. For example, consulting businesses give expert advice to other professionals, in exchange for a fee for the amount of time the constultant spent on the problem. The delivery of 'expert advice' is generally in the form of a document (which could be anything from a report to hundreds of drawings) and can range from an hours advice to thousands of hours work on a substantial project. Before a project commences, the client and consultant must agree on a fee for the expert advice, or a fee structure. The nature of how a fee is structured can be as creative as the engaged parties wish, however a couple of simple examples are as follows: the client agrees to pay for the consultants time by the hour until the task is completed, or the consultant provides a fixed fee to complete the project in full, regardless of their hours spent. The nature of the fee structure brings an element of financial risk to one party which must be mitigated.

This thesis focuses on the risk a consulting company undertakes when setting a fixed fee for a project before their role has commenced. Clients commonly ask for competitive fixed quotes from several consultants before settling with one or negotiating further. The way a consulting manager decides on their fixed fee varies greatly from industry to industry and company to company. Typically, consulting manager has extensive experience in the type of project he is quoting and can use a combination of intuition, ratios of the entire project cost, and looking up a couple of past similar projects. Such a method often results in a wide range of profitabilities for projects depending on the ability of the manager to predict the amount of time the project will demand from the consulting team. This thesis examines a case study consulting company that has recorded internal historical project performance data since 2002. The study tests whether a statistical predictive model based on past project data can predict the profitability of a project based on characteristics available at the beginning of a project. This thesis asks further whether the prediction can impact a consulting manager's determination of a fixed fee.

@@

In this chapter the research motivations are explained along with the goals of the project and the specific research hypothesis. Finally the contributions of this body of work are outlined along with the structure of the thesis.

@@

Project decision makers have struggled with forecasting project costs accurately for decades, across many industries. A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% (Flyvbjerg, 2006). A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average (Flyvbjerg and Budzier, 2011). These statistics cast a concerning glimpse at the financial story playing out behind the scenes of these projects which can cause job losses, ruin private businesses, and churn through government balancesheets. Complex long term projects are necessary to produce  railways, houses, websites, and movies. This research focuses on the financial risk that is transferred to consulting companies who are enlisted to complete portions of these complex projects. The risk is transferred by requiring a fixed flat fee from the consultant, meaning they commit to complete their role for a nominated price. In a similar way to infrastructure and IT projects, consultants are susceptible to underestimating the amount of their time that will be spent in their role, and must wear the cost of extra employee hours spent. For privately owned businesses, mitigating this risk would increase the chances of the business surviving as well as growing and succeeding.

The ramifications of unprofitable projects in a private business are multifold. Employee morale deteoriates and employees may view themselves as incompetent if they are responsible for the project's delivery. Employees also experience stress attempting to complete the project within a disappearing budget and may produce lower quality work in exchange for speed. This is unfortunate if in reality, the project fee was under estimated. Unprofitable projects also limit a business' opportunity to invest money into marketing or staff training, which would improve a business' standing. The financial risk taken on fixed flat fee projects may also discourage business' from taking different risks with better potential pay-offs, such as expanding the business into a new area. Logically, predicting fixed fee project costs more accurately could positively affect a business' growth opportunities and staff morale.

**Why its Not Easily Solvable:**

The challenge for consultants to reduce the risk of a project going over budget, in terms of hours spent, is not easily solved. Complex projects are always different from each other, and similar work may take significantly different amounts of time. For example, an ad filmed in a remote location for chocolate will cost significantly more than an ad filmed on an exisitng set, even though the product structure is almost identical. Therefore, a fixed product price, which you could apply to products such as appliances, is inappropriate for film projects. One option is charging for consultants' time and the hiring of materials by the hour; a zero risk fee structure. However in many industries this is not competitive enough. Other consultancies may be willing to offer a fixed flat fee for the same work, reducing financial risk for the client, and luring the client to them. Another fee strategy may be to track worker hours, and stop work once the consulting fee has been depleted. Then, negotiate further fees, or variations, with the client and persuade them to accept. This also satisfactorily reduces consultant financial risk, and in some cases, depending on the nature of the work or uncontrollable changes, this method can be effective. However, if the project has been following a predictable process but the hours are nevertheless inflating beyond expectation (a common situation), consultants may prefer to wear losses or endure very marginable profits to preserve the relationship with their client. Consultant's may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry over creating financial friction. Clearly, there is no straightforward solution to reducing the financial risk taken on by consultant's who offer fixed fees in complex projects. In some indusries, fixed flat fee structures are expected and business' must conform to survive.

**Current Practise:**

Limited research is available on the cost estimation techniques performed by industry for complex projects, however two surveys were performed in 2000: one on construction projects, the other on IT projects. The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practises. They found that the main method for cost estimation was breaking the project into detailed parts and adding up the cost of each item. The next two most popular methods were 'comparison with similar projects based on documented facts', and 'comparison with similar projects based on personal experience'. These can all be classified as experience based models [@Akintoye2000]. A survey by Moores and Edwards in 1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project. In the same survey, 91% of software companies cited cost estimation as a problem. Although limited recent literature is available on industry cost estimation practises, historically it can be observed that for complex construction and software projects, cost predictions are chiefly made with a mixture of experience, project comparisons, and an analysis of the project details and tasks. It is also acknowledged that cost estimation is a problem and hypothesised that inexperience and lack of time are chief contributors [@Akintoye2000].

@Lovallo2003 tackled the psychology behind the high failure rates of executives in predicting costs of projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development. Their theory stems from Kahneman's work on decision making that won him the nobel prize for economics in 2002. His research argues that humans' natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project. A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplay or ignore the risk of problems that are out of the team's control. It is for this reason that it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee. Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003]. @Lovallo2003 call the practise of analysing a project based on project details and its unique complexities the 'inside view'. Research has shown that if people are forced to make predictions about their skills after being exposed to an 'outside view', their predictions are significantly more accurate. This outside view can be applied to complex projects, and involves ignoring the details of the current project and instead analysing the outcomes of several projects of a similar description. Mobilising the outside view in this way is called reference class forecasting [@Lovallo2003]. One method @Lovallo2003 recommend is to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost. The correlation for the current project can then be estimated via a statistical model which is used to adjust the forecasted cost made by detailed analysis (the inside view) [@Flyvbjerg2011]. The traditional way to think about a project is to focus on the project details. Even though gathering data from similar projects could significantly improve a cost prediction, it is rarely thought of by managers [@Flyvbjerg2011].

The implementation of reference class forecasting began in project management for the first time in 2004, was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, and Australia. An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].

@@

The theory of reference class forecasting was first clearly published in 2003, however the idea of predicting project costs from previous data has been the subject of research for many years before that and continues to this day. Although research results have often been promising, industry uptake of the idea is not yet established.

### Overview of Past Reference Class Forecasting Methods and relevant Industry Uptake

In the following section, an overview of the historical attempts to integrate reference class forecasting will be presented followed by published theories for the lack of industry uptake.

@@

The p-value for these models and the afore-mentioned null hypothesis is `r t.test(results.rf2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`. The null hypothesis is rejected which contrasts the parallel ANOVA test. On the right hand side, of the graph above, the RMSE is improved by almost $0.20 but also worsens by over $0.10 per dollar. A density plot of the RMSE for the models is shown below:

@@

The two regression models predicting return per dollar resulted in high mean RMSE's ranging from $`r mean(results.rf.core$final.RMSE) %>% round(2)` to $`r mean(results.A2$final.RMSE) %>% round(2)`. Because of the poor results, further development of regression models was postponed in favour of investiation of binary classification models. The classification is a simpler task than regression because it requires less precision.

@@

Using this logic, if the threshold was 0, all projects were rejected and total profit would equal 0. If the threshold was 1, all projects were accepted and the total profits would be the same as historical figures. Probability results from the two blended models as well as three individual models were tested and because the models require a training and testing set, the process was repeated 100 times with 20 x 5 fold cross validation models to achieve a statistical power of 0.8. The plot below illustrates the results of each model at each threshold with a local polynomial regression line of best fit (loess) and a 95% confidence interval around the mean.

```{r profit_curve, fig.depth = 10}
finance.plot = readRDS("plot_finance.rds" )

@@

The blended models significantly (?) outperformed individual models which was expected given the two blended models' AUC results were significantly higher than the original logistic regression results (which were in line with random forests' and boosted trees' results). The simple average of logistic regression, random forests, and boosted trees results produced a better profit curve than a logistic regression of the same three variables.  The model produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t`. To put this in context, if the company's yearly profits were \$1,000,000, the mean yearly increased profit would be \$`r (best.mean-100)/100*1000000` by rejecting all jobs with a probability higher than `r best.t`. In the data, projects assigned a probability higher than `r best.t` represent only `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of all projects. The improvements in profit sound promising, however it is important to consider possible limitations in the data which is discussed in the next section.

@@

One of the unknown factors in translating this research into business practises is how well users can guess the project variables. The data used to develop the model was obtained from the company's records of how projects played out and there are no records of initially expected values of variables. For example, if a project lasted 2 years and it was initially expected it would only take 1.5 years, the only data the company retains is that the project lasted two years. This begs the question, how well could a user predict the variables before beginning a project? Uncertainty has been managed to a degree by binning variables such as 'total amount invoiced' and 'timespan' to 5 or 6 broad categories (i.e. "1 day - 3 weeks", "3 weeks - 2.5 months", etc). However, two remaining categories: team size and percent of project completed by professional level employees were left as continuous variables. It would be interesting to test how well business managers could guess these four variables at the start of a project.

>>>>

Cook, A., Wu, P., & Mengersen, K. (2015, September). Machine Learning and Visual Analytics for Consulting Business Decision Support. In Big Data Visual Analytics (BDVA), 2015 (pp. 1-2). IEEE.

@@

Predicting how long it will take a person to do something is a notoriously tricky task. People make mental calculations of these sort every day- guessing how long it will take to drive to meet someone on time, how many tasks you can check off in a day, or how long it will take to cook a dish. As most people have experienced, these estimates can be particularly inaccurate if the individual is attempting the task for the first time. These errors are trivial in everyday activities, but they have more meaning for businesses that sell their time as their chief revenue. Consulting businesses give expert advice to other professionals, in exchange for a fee for the amount of time a constultant spent on the problem. The delivery of 'expert advice' is generally in the form of a document (which could be anything from a report to hundreds of drawings) and can range from an hours advice to thousands of hours work on a substantial project. Before a project commences, the client and consultant must agree on a fee for the expert advice, or at least a fee structure. The nature of how a fee is formulated can be as creative as the engaged parties wish, however a couple of simple examples are as follows: the client agrees to pay for the consultants time by the hour until the task is complete, or the consultant provides a fixed fee to complete the project in full, regardless of their hours spent. Both contracts assign financial risk to one party which they must then mitigate to the best of their abilities.

This thesis focuses on the risk taken upon by a consulting company when agreeing to a fixed fee. Clients commonly collect fixed quotes from several consultants before settling with one or negotiating further so fees must be competitive. The way a consulting manager calculates their fixed fee varies from industry to industry and even company to company. Typically a consulting manager has experience in the type of project he is quoting and can use a combination of intuition, rules of thumb (such as a fraction of the entire project cost), and comparison to past projects. Such methods can result in a range of profit margins which ultimately depend on how well the manager predicts the amount of time a project demands. This thesis examines the problem of cost prediction via a case study on a company's internal project performance data for the previous twelve years. The study tests whether a statistical model based on historical project data can predict the profitability of a project using characteristics available at the beginning of a project. Furthermore, this thesis investigates how the model's predictions could impact a consulting manager's determination of a fixed fee and the resulting impact on the company's bottom line.

@@

In this chapter the research motivations are explained along with the goals of the project and the specific research hypotheses. Finally the contributions of this body of work are outlined along with the structure of the thesis.

@@

Project managers have struggled with forecasting project costs accurately for decades, across many industries. A study on large-scale infrastructure projects over the past seventy years revealed that cost forecasts consistently underestimated the cost of rail projects by 44.7%, the cost of bridge and tunnel projects by 33.8% and the cost of road projects by 20.4% [@Flyvbjerg2007]. A study across 1471 IT projects showed that 27% of projects ran over budget, and one in 6 of those projects were more than 200% over budget on average [@Flyvbjerg2011]. These statistics reveal a concerning glimpse at projects' hidden financial stories which can cause job losses, ruin private businesses, and churn through government balancesheets. Complex long term projects are necessary for modern society producing infrastructure, houses, websites, and movies. This research focuses on the companies who complete portions of these complex projects and the financial risk that is contractually transferred to them. Specifically, the risk of the project owner is transferred to their sub-consultant by mandating fixed flat fee structures to complete a given amount of work. The subconsultant commits to completing their role in a project for a nominated price, regardless of whether the work takes less or more time than intially anticipated. In a similar way to infrastructure and IT project budgets, consultants are susceptible to underestimating the amount of time a project will demand, and eventually wear the cost of extra employee hours spent. For privately owned businesses, mitigating this risk would increase the chances of the business surviving as well as growing and succeeding.

The ramifications of unprofitable projects are multifold. Employee morale deteoriates and employees may view themselves as incompetent if they are responsible for the project's delivery. Employees also experience stress attempting to complete the project within a disappearing budget and may produce lower quality work in exchange for speed. This experience is unfortunate if in reality, the project fee was under-estimated. Unprofitable projects also limit a business' opportunity to invest money into marketing or staff training, which would foster growth and futher profits. The financial risk taken on fixed flat fee projects may also discourage business' from taking other, more thought-out risks, such as expanding the business into a new area. Logically, predicting fixed fee project costs more accurately could positively affect business growth opportunities and staff morale.

**Why It's Not Easily Solvable:**

The challenge for consultants to reduce budget blowouts, in terms of hours spent, is not easily solved. Complex projects always differ from one another, and similar work may demand significantly different amounts of time. For example, an ad filmed in a remote location will cost significantly more than an ad filmed on an existing set, even though the effect of the final advert could be identical. Therefore, a constant fixed price to produce an advert is clearly inappropriate for film projects, but some kind of fixed price is often demanded by the client before a project is awarded. Charging clients by the hour is not competitive enough in many industries. Other consultancies are often willing to offer a fixed flat fee for the same work, luring the client with reduced financial risk on their part.

Another strategy to mitigate risks could be to track employee hours, and stop work once the fixed consulting fee has been depleted. Then, negotiate further fees, or variations, with the client and persuade them to accept. This satisfactorily reduces consultant financial risk, and in some cases, depending on the nature of the work or uncontrollable changes, this method can be effective. However, if the project has been following a predictable process but the hours are nevertheless inflating beyond expectation (a common situation), consultants may prefer to wear losses or endure marginable profits to preserve the relationship with their client. Consultants may believe it is in their best interest to maintain a reliable, trustworthy reputation in their industry above creating financial friction. Clearly, there is not a straightforward contractual solution to reducing the financial risk taken on by consultant's who offer fixed fees in complex projects. In many industries, the reality is fixed flat fee structures are expected and business' must conform to survive.

**Current Practice:**

Limited research is available on complex-project cost estimation techniques performed by industry. However, two surveys are available from 2000: one on construction projects, the other on IT projects. The construction cost estimation study by @Akintoye2000 surveyed 84 UK construction contractors, ranging from small to medium to large, about their estimating practises. They found that the main method for cost estimation was breaking the project into detailed parts and adding up the cost of each item. The next two most popular methods were 'comparison with similar projects based on documented facts', and 'comparison with similar projects based on personal experience'. These can all be classified as experience based models [@Akintoye2000].

A survey by Moores and Edwards in 1992 of 54 software developing companies found that detailed project planning tools were used by most companies as opposed to cost estimation tools, suggesting that projects were priced based on an analysis of a detailed breakdown of tasks within a project. In the same survey, 91% of software companies cited cost estimation as a problem. Although limited recent literature is available on industry cost estimation practises, historically it can be observed that for complex construction and software projects, cost predictions are chiefly made with a mixture of experience, project comparisons, and an analysis of the project details and tasks. It is also acknowledged that cost estimation is a problem and hypothesised that inexperience and lack of time are chief contributors [@Akintoye2000].

**Cost Estimation and Human Nature**

@Lovallo2003 tackled the psychology behind the high failure rates of executives in predicting costs of projects such as manufacturing plant construction, mergers and acquisitions, large infrastructure and software development. Their theory stems from Kahneman's work on decision-making that won him the nobel prize for economics in 2002. His research argues that a person's natural optimistic view of their own skills leads to consistent underestimation of the time and risks involved in a project. A manager optimistically sees challenges in a project as something that can be overcome by the team's high skill level, and downplays or ignores the risk of problems that are out of the team's control. It is for this reason that it does not matter if the project is broken down to the highest level of detail for cost prediction, all complex projects are at risk of encountering a multitude of problems that the manager could never foresee. Each problem has a low chance of occurring, but in combination the risk is much greater [@Lovallo2003].

@Lovallo2003 call the practise of analysing a project based on project details and its unique complexities the 'inside view'. This constrasts with an 'outside view' where a subject is analysed by comparison to other subjects and disregarding specific subject details. For example, research has shown that if people make predictions about their skills after being exposed to an 'outside view' (a summary of other people's skills), their predictions are significantly more accurate. According to @Locallo2003, this 'outside view' can be applied to complex projects, and stipulates that the details of the current project are ignored in favour of analysing outcomes of several similar projects. Mobilising the outside view in this way is called reference class forecasting [@Lovallo2003]. One method @Lovallo2003 recommend is to obtain correlation statistics from past similar projects - the correlation between the forecast cost and the actual cost. The correlation for the current project can then be estimated via a statistical model which is then used to adjust the forecasted cost made by detailed analysis (the inside view) [@Flyvbjerg2011]. Managers rarely analyse data from similar projects even though it could significantly improve cost predictions [@Flyvbjerg2011]. This tendency led to a formalised introduction of 'outside view' methods into some institutions.

The implementation of reference class forecasting began in project management for the first time in 2004. It was endorsed by the American Planning Association in 2005, and is now used in some governments and private companies in Europe, South Africa, and Australia. An example of the type of output from a reference class statistical model is a plot showing the relationship between the acceptable chance of cost overrun and the required uplift to the original forecasted cost [@Flyvbjerg2011].

@@

The theory of reference class forecasting was first definitively published in 2003, however the idea of predicting project costs from previous data has been the subject of research for many preceding years and continues to this day. Although research results have often been promising, industry uptake of the idea is not yet established.

### Further Reference Class Forecasting Models and Industry Uptake

In the following section, an overview of several historical attempts to integrate reference class forecasting will be presented followed by published theories for the lack of industry uptake.

@@

The p-value for these models and the afore-mentioned null hypothesis is `r t.test(results.rf2$diff, alternative = "greater", mu = 0)$p.value %>% format(digits = 2)`. The null hypothesis is rejected which contrasts the parallel ANOVA test. On the right hand side, of the graph above, the RMSE is improved by almost \$0.20 but also worsens by over \$0.10 per dollar. A density plot of the RMSE for the models is shown below:

@@

The two regression models predicting return per dollar resulted in high mean RMSE's ranging from $`r mean(results.rf.core$final.RMSE) %>% round(2)` to \$`r mean(results.A2$final.RMSE) %>% round(2)`. Because of the poor results, further development of regression models was postponed in favour of investiation of binary classification models. The classification is a simpler task than regression because it requires less precision.

@@

```{r power_profit_curve}

# How to achieve statistical power of 0.8?
finance.plot = readRDS("report_data/plot_finance.rds" )
fin.summary.1 = finance.plot %>% group_by(threshold, method) %>% 
        summarise_each(funs(mean_profit_ratio = round(mean(.), 1), 
                            sd_prof_ratio = round(sd(.), 1)),
                       profit.ratio)

#for each method find max mean 
best.fin.1 = fin.summary.1 %>% group_by(method) %>% 
        filter(mean_profit_ratio == max(mean_profit_ratio)) %>% 
        arrange(desc(mean_profit_ratio))

#two sample power test between each method and its % data points at its threshold
finance.plot$threshold <- as.factor(finance.plot$threshold)
power.fin = finance.plot %>% filter(threshold == '0.55' & method == "average"|
                                    threshold == '0.6' & method == "orig.boost"|
                                    threshold == '0.75' & method == "orig.log"|
                                    threshold == '0.55' & method == "orig.rf"|
                                    threshold == '0.6' & method == "simp.log")

# test each sample against "average" because it's the best
# n.samples function is defined in 'draft.rmd'
index = c("orig.boost","orig.log", "orig.rf","simp.log")
ns = NULL
for(i in seq_along(index)){
        temp = n.samples(data = list(power.fin %>% filter(method == index[i]) %>% .$profit.ratio, power.fin %>% filter(method == 'average') %>% .$profit.ratio), type.calc = 'two.sample', alt = 'two.sided')
        ns[i] = temp
}
ns

# orig.boost needs 6, orig.log needs 4, orig.rf needs 6 and simp.log needs 64 : ) do 100!

Using this logic, if the threshold was 0, all projects were rejected and total profit would equal 0. If the threshold was 1, all projects were accepted and the total profits would be the same as historical figures. Probability results from the two blended models as well as three individual models were tested and because the models require a training and testing set, the process was repeated 100 times with 20 x 5 fold cross validation models to achieve a statistical power of 0.8 (calculated after an initial run of 10 models). The plot below illustrates the results of each model at each threshold with a local polynomial regression line of best fit (loess) and a 95% confidence interval around the mean.

```{r profit_curve, fig.depth = 10}

finance.plot = readRDS("report_data/plot_finance.rds" )

@@

The blended models significantly (?) outperformed individual models which was expected given the two blended models' AUC results were significantly higher than the original logistic regression results (which were in line with random forests' and boosted trees' results). The simple average of logistic regression, random forests, and boosted trees results produced a better profit curve than a logistic regression of the same three variables.  The model produced a `r best.mean-100`% increase in profit by rejecting jobs with a probability of loss over `r best.t`. To put this in context, if the company's yearly profits were \$1,000,000, the mean yearly increased profit would be \$`r (best.mean-100)/100*1000000` by rejecting all jobs with a probability higher than `r best.t`. In the data, projects assigned a probability higher than `r best.t` represent only `r ((simp.l %>% filter(mean.iter>best.t) %>% nrow)/nrow(simp.l)*100) %>% round(1)`% of all projects.

Alternative strategies exist for how to deal with projects that are marked with a probability of loss greater than `r best.t`. For example, contracts for these projects could be changed to hourly rate contracts without exceptions. This means that each hour booked on the project will definitely be charged to the client for each stage of the project. Another example would be to increase fixed fees for projects above the threshold by a nominal percentage. However, this method seems less practical as the case study projects above `r best.t` would need to increase fees by 46% in order to generate a 15% profit. A proposed fee increase of that magnitude may not be as acceptable as an hourly rate contract depending on the project and client. Contracts could be adjusted in many other creative ways to reduce risk in high probability projects, and the threshold of `r best.t` was optimised based on the accept/reject strategy. Different courses of actions could be developed for specific clients or project sizes via analysis of threshold scenarios.

@@

The improvements in profit sound promising, however it is important to consider possible limitations in the data which is discussed in the next section. One of the unknown factors in translating this research into business practises is how well users can guess the project variables. The data used to develop the model was obtained from the company's records of how projects played out and there are no records of initially expected values of variables. For example, if a project lasted 2 years and it was initially expected it would only take 1.5 years, the only data the company retains is that the project lasted two years. This begs the question, how well could a user predict the variables before beginning a project? Uncertainty has been managed to a degree by binning variables such as 'total amount invoiced' and 'timespan' to 5 or 6 broad categories (i.e. "1 day - 3 weeks", "3 weeks - 2.5 months", etc). However, two remaining categories: team size and percent of project completed by professional level employees were left as continuous variables. It would be interesting to test how well business managers could guess these four variables at the start of a project.
